Automated detection of data fabrication using statistical tools

### 2018-05-03 submission/redraft.rmd
Add paragraphs

```{r echo = FALSE}
source('../functions/digit_analysis.R')
source('../functions/std_var.R')
source('../functions/fisher_method.R')

@@

Effect sizes can be reported in research reports in various ways. 
For example, the most commons ways effect sizes are reported in papers are as a standardized mean difference (e.g., $d$), as an explained variance (e.g., $R^2$), or as a test statistic. 
A test statistic is also a measure of effect size, albeit in a not directly interpretable form. 
A test result such as $t(
`r df`)=`r tval`$ corresponds to $d=`r round(2*tval / sqrt(df), 3)` and $r=`r round((tval^2 / df) / ((tval^2 / df) + 1), 3)` [@10.1525/collabra.71]. 
These effect sizes can readily be recomputed based on data extracted with `statcheck` across thousands of results [@10.3758/s13428-015-0664-2;@10.3390/data1030014].

@@

@@

@@

@@

>>>>

```{r echo = FALSE}
library(ggplot2)

source('../functions/digit_analysis.R')
source('../functions/std_var.R')
source('../functions/fisher_method.R')

ml_file <- '../data/study_01/ml_summary_stats.csv'
qualtrics_file <- '../data/study_01/qualtrics_processed.csv'
dat_summary_file <- '../data/study_01/study1_res.csv'
file <- '../data/study_01/raw_summary_results_fabrication_qualtrics.csv'
res_file <- '../data/study_01/qualtrics_processed.csv'
ml_dat_file <- '../data/study_01/anchoring_ml/chjh ml1_anchoring cleaned.sav'
summary_stat_file <- '../data/study_01/ml_summary_stats.csv'
ml3_dat_file <- '../data/study_02/study_02-ml3_stroop/StroopCleanSet.csv'
pdf_file <- '../archive/gender_interaction.pdf'

# Set the number of iterations to use in calculations
iter <- 100000

@@

Effect sizes can be reported in research reports in various ways. 
For example, the most commons ways effect sizes are reported in papers are as a standardized mean difference (e.g., $d$), as an explained variance (e.g., $R^2$), or as a test statistic. 
A test statistic is also a measure of effect size, albeit in a not directly interpretable form. 
A test result such as $t(
`r df`)=`r tval`$ corresponds to d=`r round(2*tval / sqrt(df), 3)` and r=`r round((tval^2 / df) / ((tval^2 / df) + 1), 3)` [@10.1525/collabra.71]. 
These effect sizes can readily be recomputed based on data extracted with `statcheck` across thousands of results [@10.3758/s13428-015-0664-2;@10.3390/data1030014].

@@

## Detecting data fabrication in raw data

@@

```{r prep study 1, echo=FALSE}
# Compute the summary statistics for Many Labs
suppressWarnings(suppressMessages(source('../functions/compute_summary_anch_ml.R')))
# Process the collected, fabricated data
suppressMessages(source('../functions/process_qualtrics_anch_01.R'))
# Concatenate, analyze, and write out all ML and qualtrics data
if(!file.exists('../data/study_01/study1_res.csv'))
{
 set.seed(123);suppressMessages(source('../functions/concatenate_analyze_01.R')) 
}

dat_summary <- read.csv(dat_summary_file)

type_index <- which(names(dat_summary) == 'type')
temp <- dat_summary[, c(type_index, 18:length(names(dat_summary)))] 
dat_summary$es_gender <- sqrt(dat_summary$es2_gender)
dat_summary$es_condition <- sqrt(dat_summary$es2_condition)
dat_summary$es_interaction <- sqrt(dat_summary$es2_interaction)

# naming just for ease of use later
dat_summary$es_p <- 1 - dat_summary$es_condition

# ROC
AUC <- apply(dat_summary[, grepl(names(dat_summary), pattern = "_p")],
   2,
   function(x)
   {
    y <- pROC::roc(response = dat_summary$type, x, direction = ">")
    return(y)
   }
)
```

@@

Figure `r fignr` shows a group-level comparison of the genuine- and fabricated $p$-values and relevant effect sizes ($r$). 
These group-level comparisons provide an overview of the differences between the genuine- and fabricated data [see also @10.1186/1471-2288-3-18]. 
These distributions indicate little group differences between genuine- and fabricated data when nonsignificant effects are inspected (i.e., gender and interaction hypotheses). 
However, there seem to be large group differences when we required subjects to fabricate significant summary statistics (i.e., condition hypothesis).

```{r ddfab_density, fig.cap="Overlay of density distributions for both genuine and fabricated data, per effect and type of result. We instructed respondents to fabricate nonsignificant summary statistics for the gender and interaction effects, and a significant effect for the condition effect.", out.width='100%', echo=FALSE}
plot_p_gender <- ggplot(dat_summary, aes(x = p_gender, fill = type)) + 
 geom_density(alpha = .3) + 
 xlab(latex2exp::TeX("$P$-value")) + 
 ylab("Density") + 
 ylim(0, 1.5) + 
 scale_fill_discrete(guide=FALSE) + 
 ggtitle('Gender')

plot_es_gender <- ggplot(dat_summary, aes(x = es_gender, fill = type)) + 
 geom_density(alpha = .3) + 
 xlab(latex2exp::TeX("Effect size ($r$)")) + 
 ylab("Density") + 
 scale_fill_discrete(guide=FALSE)

# Condition
dat_summary_tmp <- dat_summary
dat_summary_tmp$p_condition <- log10(dat_summary_tmp$p_condition)
plot_p_condition <- ggplot(dat_summary_tmp, aes(x = p_condition, fill = type)) +
 geom_density(alpha = .3) +
 xlab(latex2exp::TeX("log10 $P$-value")) +
 ylab("Density") +
 scale_fill_discrete(labels = c("Genuine","Fabricated")) + 
 theme(legend.position="top") +
 ggtitle('Condition')

plot_p_condition$labels$fill <- ""

plot_es_condition <- ggplot(dat_summary, aes(x = es_condition, fill = type)) + 
 geom_density(alpha = .3) + 
 xlim(0, 1) + 
 xlab(latex2exp::TeX("Effect size ($r$)")) + 
 ylab("Density") + 
 scale_fill_discrete(guide=FALSE)

# Interaction
plot_p_interaction <- ggplot(dat_summary, aes(x = p_interaction, fill = type)) + 
 geom_density(alpha = .3) + 
 xlab(latex2exp::TeX("$P$-value")) + 
 ylab("Density") + 
 scale_fill_discrete(guide=FALSE) +
 ylim(0, 1.5) +
 ggtitle('Interaction')

plot_es_interaction <- ggplot(dat_summary, aes(x = es_interaction, fill = type)) + 
 geom_density(alpha = .3) + 
 xlab(latex2exp::TeX("Effect size ($r$)")) + 
 ylab("Density") + 
 scale_fill_discrete(guide=FALSE)

# Plot
gridExtra::grid.arrange(plot_p_gender,
       plot_p_condition,
       plot_p_interaction,
       plot_es_gender,
       plot_es_condition,
       plot_es_interaction,
       ncol = 3)

Using nonsignificant $p$-values to detect data fabrication is hardly better than chance level in the current sample. 
We asked researchers to fabricate data for nonsignificant effect sizes across four anchoring studies, thinking they might be unable to produce uniformly distributed $p$-values due to misunderstanding of what a $p$-value means [@;@]. 
<!-- add p-value interpretation from ej, sijtsma p value paper -->
Our results indicate that the $p$-values for the gender effect and interaction effect perform only slightly above chance level for detecting fabricated data, $AUROC=`r round(AUC$gender_fish_p$auc[1], 3)`$ and $AUROC=`r round(AUC$interaction_fish_p$auc[1], 3)`$, respectively. 
In other words, widespread detection based on nonsignificant $p$-values does not seem promising.

```{r echo = FALSE}
selector <- grepl(names(AUC), pattern = '(variance_|maxmin_)')
selector_sd <- grepl(names(AUC), pattern = '(variance_)')
selector_maxmin <- grepl(names(AUC), pattern = '(maxmin_)')
```

We performed a total of `r sum(selector)` variance analyses across the fabricated standard deviations for the four anchoring studies, with `r sum(selector_sd)` using the $SD_z$ measure and `r sum(selector_maxmin)` using the $max-min_z$ measure. More specifically, we analyzed the standard deviations across all studies combined (assuming homogeneity across conditions), across all studies (assuming heterogeneity across anchoring conditions), within each study, and within each anchoring condition. We computed the AUROC value with the directional assumption that genuine data would show more variation than fabricated data. Results are presented in Table `r tabnr`.

```{r table_variance, echo = FALSE}
dfvar <- data.frame(Method = c("Homogeneous, all studies combined",
                "Homogeneous, study 1",
                "Homogeneous, study 2",
                "Homogeneous, study 3",
                "Homogeneous, study 4",
                "Heterogeneous, all studies combined",
                "Heterogeneous study 1, low anchor condition",
                "Heterogeneous study 1, high anchor condition",
                "Heterogeneous study 2, low anchor condition",
                "Heterogeneous study 2, high anchor condition",
                "Heterogeneous study 3, low anchor condition",
                "Heterogeneous study 3, high anchor condition",
                "Heterogeneous study 4, low anchor condition",
                "Heterogeneous study 4, high anchor condition"),
          SD_z = c(
            AUC$variance_sd_p_overall_homo$auc,
            AUC$variance_sd_p_study1$auc,
            AUC$variance_sd_p_study2$auc,
            AUC$variance_sd_p_study3$auc,
            AUC$variance_sd_p_study4$auc,
            AUC$variance_sd_p_overall_hetero$auc,
            AUC$variance_sd_p_study1_low$auc,
            AUC$variance_sd_p_study1_high$auc,
            AUC$variance_sd_p_study2_low$auc,
            AUC$variance_sd_p_study2_high$auc,
            AUC$variance_sd_p_study3_low$auc,
            AUC$variance_sd_p_study3_high$auc,
            AUC$variance_sd_p_study4_low$auc,
            AUC$variance_sd_p_study4_high$auc
            ),
          max_min_z = c(
            AUC$maxmin_p_overall_homo$auc,
            AUC$maxmin_p_study1$auc,
            AUC$maxmin_p_study2$auc,
            AUC$maxmin_p_study3$auc,
            AUC$maxmin_p_study4$auc,
            AUC$maxmin_p_overall_hetero$auc,
            AUC$maxmin_p_study1_low$auc,
            AUC$maxmin_p_study1_high$auc,
            AUC$maxmin_p_study2_low$auc,
            AUC$maxmin_p_study2_high$auc,
            AUC$maxmin_p_study3_low$auc,
            AUC$maxmin_p_study3_high$auc,
            AUC$maxmin_p_study4_low$auc,
            AUC$maxmin_p_study4_high$auc
            ))

names(dfvar) <- c('Method', 'AUROC $SD_z$', 'AUROC $max-min_z$')
knitr::kable(dfvar, row.names = FALSE, digits = 3, caption = "_Table XX._ Diagnosticity of using variance analyses to detect data fabrication, depicted with the AUROC-value. We assumed fabricated results would show ")
tabnr <- tabnr + 1
```

Results indicate that (1) heterogeneity of population variances directly affect the effectiveness of detecting data fabrication, (2) the different operationalizations of variance analyses show minimal differences, and (3) that the combined analysis of standard deviations across anchoring conditions in the four studies is a preferred method. 
When comparing the combined variance analyses, we see that the homogeneity assumption drastically decreases the performance ($SD_z$; homogeneity: $AUROC=`r round(AUC$variance_sd_p_overall_homo$auc, 3))`$, heterogeneity: AUROC=`r round(AUC$variance_sd_p_overall_hetero$auc, 3))`). 
This assumption has not previously been explicated [@] and we also did not preregister analyses that took into account this assumption. 
Moreover, the maximum absolute difference between the $SD_z$ and $max-min_z$ operationalizations is `r round(max(abs(dfvar$SD_z - dfvar$max_min_z)), 3)` and there is no reason to deviate from the original $SD_z$ operationalization as proposed by @. 
Lastly, we see that variance analyses separated per study or anchoring condition within a study are quite variable, which suggests that the combined analysis is preferred because selecting a priori selection of a specific subset is unfeasible in practice.

By simply comparing effect sizes of the main anchoring effect, results indicate that this is a parsiminious and effective way of detecting data fabrication. More specifically, given a genuine- and fabricated anchoring effect size, there is approximately a 75% chance that the larger effect size is the fabricated one (i.e., $AUROC=`r round(AUC$es_p$auc, 3)`$). Comparing this to the best performing variance analysis (i.e., $AUROC=`r round(AUC$variance_sd_p_overall_hetero$auc, 3)`$), performance is similar but considerably easier in practice.

### Combination of p-value and variance analysis

@@

### Digit analysis

#### Newcomb-Benford Law

#### Terminal digit analysis

#### Multivariate associations
